---
title: "Using the MicroBioMap package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the MicroBioMap package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

## Overview

## Getting started

### Installation

```{r eval=FALSE}
library(BiocManager)
BiocManager::install('seandavi/MicroBioMap')
```

## Basic usage

```{r message=FALSE}
library(MicroBioMap)
# this operation requires about 4GB of RAM
cpd <- get_compendium()
```

```{r}
cpd
```

### Sample metadata


The `colData` slot of the `cpd` object contains the information about samples.
We can examine the names of the available metadata:

```{r}
names(colData(cpd))
```

We can look at the samples by available regions:

* Region
    ```{r}
    table(colData(cpd)$region)
    ```

* Library strategy
    ```{r}
    table(colData(cpd)$library_strategy)
    ```

* Instrument
    ```{r}
    table(colData(cpd)$instrument)
    ```

* Country
    ```{r}
    head(sort(table(colData(cpd)$country), decreasing = TRUE))
    ```


### Microbe (row) metadata

Each row in `cpd` represents one microbe. The `rowData` contains a data.frame-
like set of metadata with columns `kingdom`, `phylum`, etc. 

```{r}
head(rowData(cpd))
```

## Use cases

- [ ] TODO

### Samples in specific Bioprojects

### Samples from a specific world region

## Read counts aggregated to *Order*

## Samples filtered by presence of microbe

